Cell Decompositions for Rank Two Quiver Grassmannians

### 2018-02-06 Grassmannian_cells.tex
Section 4 edits

It will be important to understand the possible subrepresentations of the truncated preprojective representations $\tilde P_{m+1}^I$.
To describe these, we introduce notation for locating lifted preprojectives as subrepresentations of our standard lifted preprojective representations.
Since we work only on two fixed components of $\widetilde{K(n)}$, the following notation will be useful in describing paths in these components.
For $k\ge1$, set
\[A_1^{(k)}:=\{(i_1,\ldots,i_k)\mid i_j\in\{1,\ldots,n\}\text{ for $1\le j\le k$}\}.\]%, i_j\ne i_{j+1}\text{ for $1\le j<l$}\}.\]
Depending on context (in particular, the parity of $m+1$ below), we will sometimes identify the word~$(i_1,\ldots,i_k)$ with the element $\alpha_{i_1}\alpha_{i_2}^{-1}\alpha_{i_3}\cdots\alpha_{i_k}^{(-1)^{k+1}}\in W_n$ and sometimes with the element $\alpha_{i_1}^{-1}\alpha_{i_2}\alpha_{i_3}^{-1}\cdots\alpha_{i_k}^{(-1)^k}\in W_n$.
In this way, to each word $\ui=(i_1,\ldots,i_k)\in A_1^{(k)}$ with $1\leq k\leq m$, we associate a preprojective subrepresentation~$\tilde P_{m+1-k,\ui}\subset\tilde P_{m+1}$ which lifts $P_{m+1-k}$.
More precisely, we obtain a sequence of preprojective subrepresentations which uniquely determines the desired inclusion:
\[\tilde P_{m+1-k,\ui}\subset\tilde P_{m+2-k,(i_1,\ldots,i_{k-1})}\subset\ldots\subset\tilde P_{m-1,(i_1,i_2)}\subset\tilde P_{m,i_1}\subset\tilde P_{m+1}.\]
Note that, although there is a translate of $\tilde P_{m+1-k}$ corresponding to each word $\ui\in A_1^{(l)}$ for $1\le l<k$, these will not be naturally equipped with a canonical inclusion to $\tilde P_{m+1}$.

To emphasize this point, consider a word $\ui\in A_1^{(k)}$ with $i_j=i_{j+1}$ for some $j$ and write $\ui'\in A_1^{(k-2)}$ for the word obtained from $\ui$ by removing the terms $i_j$ and $i_{j+1}$.
Then the representations $\tilde P_{m+1-k,\ui}$ and $\tilde P_{m+1-k,\ui'}$ are in fact equal, however $\tilde P_{m+1-k,\ui}$ is naturally identified as a subrepresentation of $\tilde P_{m+1}$ while $\tilde P_{m+1-k,\ui'}$ is not.
Indeed, by considering the support quiver of $\tilde P_{3,i}$ from Example~\ref{ex:lifted preprojectives}, we see that each $\tilde P_{2,(i,i)}$ is just a copy of $\tilde P_2$ when viewed as representations of $\widetilde{K(n)}$, however these provide distinct subrepresentations of $\tilde P_4$.
\begin{lemma}
  \label{le:subrep}
  For $m\ge1$ and $I\subsetneq\{1,\ldots,n\}$, all non-trivial proper subrepresentations of $\tilde P_{m+1}^I$ are direct sums of preprojective representations.	
\end{lemma}
\begin{proof}
  By Lemma~\ref{le:homdecomposition}, the projected representation $G(\tilde P_{m+1}^I)$ is a truncated preprojective of $K(n)$.
  Then Lemma~\ref{le:projective subrepresentations} shows that all subrepresentations of $G(\tilde P_{m+1}^I)$ are direct sums of preprojective representations of $K(n)$.
  But any subrepresentation of $\tilde P_{m+1}^I$ also projects to a subrepresentation of $G(\tilde P_{m+1}^I)$.
  Since, following Lemma~\ref{le:lifts of transjectives}, any lift of a preprojective of $K(n)$ will be a preprojective representation of $\widetilde{K(n)}$, this gives the result.
\end{proof}

@@

For the notation of the following result, we refer the reader to the discussion before Lemma~\ref{le:subrep}.
\begin{lemma}
  \label{le:special subrepresentations}
  Consider a nonempty subset $I\subsetneq\{1,\ldots,n\}$ and fix an element $j\in I$. 
  \begin{enumerate}
    \item For $m\ge2$, we have $\Hom(\tilde P_{m,j},\tau\tilde P_{m+1}^I)\cong\CC$.
      Moreover, the kernel of a nonzero morphism $\tilde P_{m,j}\to\tau\tilde P_{m+1}^I$ is the following representation 
      \[\tilde P_m(I,j):=
        \begin{cases}
          \bigoplus_{\substack{1\leq i\leq n\\i\neq j}}\tilde P_{m-1,(j,i)}\oplus \bigoplus_{i\in I, i\ne j}\tilde P_{m-2,(j,j,i)} & \text{ if $m\geq 3$;}\\
          \bigoplus_{\substack{1\leq i\leq n\\i\neq j}}\tilde P_{1,(j,i)} & \text{ if $m=2$.}
        \end{cases}\]
    \item For $m\ge3$, any nonzero morphism $\tilde P_{m,j}\to\tau\tilde P_{m+1}^I$ is surjective.
  \end{enumerate}
\end{lemma}
\begin{proof}
  The first claim of part (1) follows immediately by applying the Auslander-Reiten formulas \cite[Theorem IV.2.13]{ass} to Lemma~\ref{le:properties}.(3).
  Indeed, this gives
  \[\dim\Hom(\tilde P_{m,j},\tau\tilde P_{m+1}^I)=\dim\Ext(\tilde P_{m+1}^I,\tilde P_{m,j})=1.\]
  We establish the final claim of part (1) directly for $m=2,3$ and then deduce the general case by applying the reflection recursions \eqref{eq:recursive covers}.
  Using the description in Example~\ref{ex:truncated lifts}, it is not hard to see that $\tau\tilde P_3^I$ is indecomposable with one-dimensional spaces at only the vertices $(1,e)$ and $(2,\alpha_i^{-1})$ for $i\in I$.
  But then for~$j\in I$, the image of the unique homomorphism $\tilde P_{2,j}\to\tau\tilde P_3^I$ is the representation with support quiver
  \[(2,\alpha_j^{-1})\xrightarrow{\alpha_j}(1,e).\]
  From the support quiver of $\tilde P_{2,j}$ given in Example~\ref{ex:lifted preprojectives}, we see that the kernel of this map is precisely $\tilde P_2(I,j)$.

@@

The following orthogonality property is a primary reason we need to lift to the universal cover of $K(n)$.
\begin{corollary}
  \label{cor:perpendicular}
  Consider a nonempty subset $I\subsetneq\{1,\ldots,n\}$ and fix an element $j\in I$.
  For $m\geq 2$, we have $\tilde P_m(I,j)\in (\tilde P_{m+1}^I)^\perp$.
\end{corollary}
\begin{proof}
  If $m\geq 3$, we consider the long exact sequence obtained when applying $\Hom(\tilde P_{m+1}^I,-)$ to the sequence
  \[\ses{\tilde P_m(I,j)}{\tilde P_{m,j}}{\tau\tilde P_{m+1}^I}.\]
  Following Lemma~\ref{le:properties}.(3), we have $\Hom(\tilde P_{m+1}^I,\tilde P_{m,j})=0$ and using the long exact sequence this immediately implies $\Hom(\tilde P_{m+1}^I,\tilde P_m(I,j))=0$.
  From Lemma~\ref{le:properties}.(3) again, we have $\Ext(\tilde P_{m+1}^I,\tilde P_{m,j})\cong\CC$.
  Using the Auslander-Reiten formulas \cite[Theorem IV.2.13]{ass} and Lemma~\ref{le:properties}.(4), we get
  \[\dim\Ext(\tilde P_{m+1}^I,\tau\tilde P_{m+1}^I)=\dim\Hom(\tilde P_{m+1}^I,\tilde P_{m+1}^I)=1.\]
  It follows that the surjective map $\Ext(\tilde P_{m+1}^I,\tilde P_{m,j})\to\Ext(\tilde P_{m+1}^I,\tau\tilde P_{m+1}^I)$ appearing in the long exact sequence is in fact an isomorphism.
  But the Auslander-Reiten formulas and Lemma~\ref{le:properties}.(4) again imply 
  \[\dim\Hom(\tilde P_{m+1}^I,\tau\tilde P_{m+1}^I)=\dim\Ext(\tilde P_{m+1}^I,\tilde P_{m+1}^I)=0.\]
  Combining with the preceding discussion, this gives $\Ext\!\big(\tilde P_{m+1}^I,\tilde P_m(I,j)\big)=0$.

@@

\subsection{Bialynicki-Birula decomposition}
The aim of Section \ref{torusaction} is to define a torus action on quiver Grassmannians which can be used to simplify the calculation of homological invariants in general.
If the quiver Grassmannian is smooth, which is for instance the case for exceptional representations by \cite{cr}, it can also be used to stratify the quiver Grassmannians using the results of Bialynicki-Birula.
More specifically, let $X$ be a smooth projective variety with a $\C^\ast$-action.
For a connected component of the fixed point set $C\subset X^{\C^\ast}$, we define its attracting set as
\[
  \mathrm{Att}(C):=\{y\in X\mid \lim_{t\to 0}t.y\in C\}
\]

The following result of Bialynicki-Birula relates the geometry of $X$ to the geometry of its torus fixed points.
\begin{theorem}\label{thm:bb}
  \cite[Section 4]{bb}
  Let $X^{\C^\ast}=\coprod_{i=1}^r C_i$ be the decomposition of the fixed point set of $X$ into connected components.
  Then $\mathrm{Att}(C_i)$ is a locally closed smooth $\C^\ast$-invariant subvariety of $X$ containing $C_i$ as a subvariety.
  Moreover, we have $X=\coprod_{i=1}^r\mathrm{Att}(C_i)$ and the natural map $\gamma_i:\mathrm{Att}(C_i)\to C_i$ is an affine vector bundle. 
  In particular, we have an equality of Euler characteristics $\chi(X)=\chi(X^{\C^\ast})$.
\end{theorem}

@@

We can consider the decomposition $X_i=\bigoplus_{(i,\chi)\in A_Q} X_{(i,\chi)}$.
We define a torus action on each $X_{(i,\chi)}$ via $t.x_{(i,\chi)}=t^{d(i,\chi)}x_{(i,\chi)}$ which can be extended linearly to each $X_q$.
For a fixed a subspace $U_i$, we can define the subspace $t.U_i$.
In general, this torus action induces no torus action on the quiver Grassmannians $\Gr_\bfe(X)$ as $t.U=(t.U_i)_{i\in Q_0}$ is no subrepresentation of $X$ for every $U\in\Gr_\bfe(X)$.
Actually, such an action has to satisfy $X_\alpha(t.U_i)\in t.U_j$ for every $\alpha:i\to j$. 

@@

  Fix an integer $c_\alpha\in\ZZ$ for every $\alpha\in Q_1$.
  If $X\in\rep Q$ can be lifted, $d:\hat Q_0\to\ZZ$ induces a torus action on $\Gr_\bfe(X)$ if we have $d(j,\chi+e_\alpha)-d(i,\chi)=c_\alpha$ for all $\alpha:i\to j$ and $\chi\in A_Q$.

@@

  Let $U\in\Gr_\bfe(X)$. Since $X$ can be lifted, we can write $X_\alpha:X_i\to X_j$ as block matrix consisting of linear maps $X_{(\alpha,\chi)}:X_{(i,\chi)}\to X_{(j,\chi+e_\alpha)}$. Let $u_{(i,\chi)}\in U_{(i,\chi)}$ and $\alpha:i\to j\in Q_1$. Then we have $X_\alpha(u_{(i,\chi)})\in U_{(j,\chi+e_\alpha)}$ because $X$ can be lifted to the universal cover. It follows
\[X_\alpha(t.u_{(i,\chi)})=X_\alpha(t^{d(i,\chi)}u_{(i,\chi)})=t^{d(i,\chi)}X_\alpha(u_{(i,\chi)})=t^{d(i,\chi)}u_{(j,\chi+e_\alpha)}=t^{-c_\alpha}t.u_{(j,\chi+e_\alpha)}\]
for some $u_{(j,\chi+e_\alpha)}\in U_{(j,\chi+e_\alpha)}$.
Now we have \[\bigoplus_{\chi\in A_Q} t.U_{(j,\chi)}=\bigoplus_{\chi\in A_Q}t^{c(j,\chi)}t.U_{(j,\chi)}\]
for every $t\in\C^\ast$ if $c(j,\chi)$ is constant along $(j,\chi)\in \{j\}\times A_Q$. This yields  
\[t.U_j=\bigoplus_{\chi\in A_Q} t.U_{(j,\chi)}=\bigoplus_{\chi\in A_Q}t^{-c_\alpha}t.U_{(j,\chi)}\]
for every $t\in \C^\ast$.

@@

\begin{lemma}
  \label{degreecondition} 
  Let $X$ be an indecomposable representation of $\hat Q$. There exists $d:\mathrm{supp} (X)\to\ZZ$ and $c_\alpha\in\mathbb N_+$ for each $\alpha\in Q_1$ such that
  \begin{enumerate}
    \item for $(i,\chi),\,(j,\chi')\in\mathrm{supp}(X)$ with $(i,\chi)\neq (j,\chi')$ we have $d(i,\chi)\ne d(j,\chi')$;
    \item for $(i,\chi),\,(j,\chi')\in\mathrm{supp}(X)$, we have $d(j,\chi')-d(i,\chi)=c_\alpha$ if and only if $\chi'=\chi+e_\alpha$.
  \end{enumerate}
\end{lemma}
\begin{proof}
  Assume that $|Q_1|=n$.
  Since $X$ is finite-dimensional and indecomposable, the support quiver $\mathrm{supp}(X)$ is a connected and finite subquiver of $\hat Q$.
  In order to prove the statement, we may assume that $(i,0)\in\mathrm{supp}(X)$ for some $i\in Q_0$.
  We fix $c_\alpha\in\NN$ and show that we can choose a degree such that the two conditions are satisfied for this choice.
  Let $K$ be the maximal length of a path in $\mathrm{supp}(X)$ starting or ending in $(i,0)$ such that the underlying graph of the path has no cycles.
  This implies that, for $(i',\chi)\in\mathrm{supp}(X)$ with $\chi=\sum_{j=1}^n d_j\alpha_j$, we have $|d_j|\leq K$.

  Set $c_{\alpha_1}=1$ and choose $c_{\alpha_j}$ recursively in such way that 
  \[c_{\alpha_k}\geq 2(K+1)\sum_{j=1}^{k-1} c_{\alpha_j}\]
  for $k=2,\ldots,n$.

  Let $f:A_Q\to\ZZ$ be the group homomorphism defined by $f(\alpha)=c_\alpha$ for all $\alpha\in Q_1$ and define $d(i',\chi):=f(\chi)$ for $(i',\chi)\in\mathrm{supp}(X)$.
  In particular, for $(i',\chi)$ with $\chi=\sum_{j=1}^n d_j\alpha_j$, we have
  \[d(i',\chi)=\sum_{j=1}^n d_jc_{\alpha_j}.\]

  First note that we have $d(j,\chi+\alpha)-d(i,\chi)=c_\alpha$.
  Now assume that $d(i,\chi)=d(j,\chi')$ which means that $d(i,\chi)=\sum_{i=1}^n d_ic_{\alpha_i}=\sum_{i=1}^n e_ic_{\alpha_i}=d(j,\chi')$ if $\chi=\sum_{i=1}^n d_ic_{\alpha_i}$ and $\chi'=\sum_{i=1}^n e_ic_{\alpha_i}$.
  This implies
  \[\sum_{i=1}^{n-1}(d_i-e_i)c_{\alpha_i}=(e_{n}-d_{n})c_{\alpha_{n}}.\]
  Since we have $|d_i-e_i|\leq 2K$ as $|d_i|,|e_i|\leq K$, we obtain 
  \[|\sum_{i=1}^{n-1}(d_i-e_i)c_{\alpha_i}|<2K\sum_{i=1}^{n-1}c_{\alpha_i}\leq c_{\alpha_n}.\]
  This inductively yields $d_i=e_i$ for $i=n,\ldots,1$ by the choice of $c_{\alpha_i}$.

  Finally, assume that $d(j,\chi')-d(i,\chi)=c_\alpha$.
  Then we have $d(j,\chi')=d(j,\chi+e_\alpha)$.
  An analogous proof as the one above shows that $\chi'=\chi+e_\alpha$.
  Note that it is a priori possible that $(j,\chi+e_\alpha)\notin\mathrm{supp}(X)$.
  Thus we chose $c_{\alpha_j}$ in such a way that it is greater than $2(K+1)\sum_{i=1}^{j-1}c_{\alpha_i}$. 
\end{proof}

>>>>

It will be important to understand the possible subrepresentations of the truncated preprojective representations $\tilde P_{m+1}^I$.

\begin{lemma}
  \label{le:subrep}
  For $m\ge1$ and $I\subsetneq\{1,\ldots,n\}$, all non-trivial proper subrepresentations of $\tilde P_{m+1}^I$ are direct sums of preprojective representations.	
\end{lemma}
\begin{proof}
  By Lemma~\ref{le:homdecomposition}, the projected representation $G(\tilde P_{m+1}^I)$ is a truncated preprojective of $K(n)$.
  Then Lemma~\ref{le:projective subrepresentations} shows that all subrepresentations of $G(\tilde P_{m+1}^I)$ are direct sums of preprojective representations of $K(n)$.
  But any subrepresentation of $\tilde P_{m+1}^I$ also projects to a subrepresentation of $G(\tilde P_{m+1}^I)$.
  Since, following Lemma~\ref{le:lifts of transjectives}, any lift of a preprojective of $K(n)$ will be a preprojective representation of $\widetilde{K(n)}$, this gives the result.
\end{proof}

@@

We now introduce notation for locating specific lifted preprojectives as subrepresentations of our standard lifted preprojective representations.
Since we work only on two fixed components of $\widetilde{K(n)}$, the following notation will be useful in describing paths in these components.
For $k\ge1$, set
\[A_1^{(k)}:=\{(i_1,\ldots,i_k)\mid i_j\in\{1,\ldots,n\}\text{ for $1\le j\le k$}\}.\]%, i_j\ne i_{j+1}\text{ for $1\le j<l$}\}.\]
Depending on context (in particular, the parity of $m+1$ below), we will sometimes identify the word~$(i_1,\ldots,i_k)$ with the element $\alpha_{i_1}\alpha_{i_2}^{-1}\alpha_{i_3}\cdots\alpha_{i_k}^{(-1)^{k+1}}\in W_n$ and sometimes with the element $\alpha_{i_1}^{-1}\alpha_{i_2}\alpha_{i_3}^{-1}\cdots\alpha_{i_k}^{(-1)^k}\in W_n$.
In this way, to each word $\ui=(i_1,\ldots,i_k)\in A_1^{(k)}$ with $1\leq k\leq m$, we associate a preprojective subrepresentation~$\tilde P_{m+1-k,\ui}\subset\tilde P_{m+1}$ which lifts $P_{m+1-k}$.
More precisely, we obtain a sequence of preprojective subrepresentations which uniquely determines the desired inclusion:
\[\tilde P_{m+1-k,\ui}\subset\tilde P_{m+2-k,(i_1,\ldots,i_{k-1})}\subset\ldots\subset\tilde P_{m-1,(i_1,i_2)}\subset\tilde P_{m,i_1}\subset\tilde P_{m+1}.\]
Note that, although there is a translate of $\tilde P_{m+1-k}$ corresponding to each word $\ui\in A_1^{(l)}$ for $1\le l<k$, these will not be naturally equipped with a canonical inclusion to $\tilde P_{m+1}$.

To emphasize this point, consider a word $\ui\in A_1^{(k)}$ with $i_j=i_{j+1}$ for some $j$ and write $\ui'\in A_1^{(k-2)}$ for the word obtained from $\ui$ by removing the terms $i_j$ and $i_{j+1}$.
Then the representations $\tilde P_{m+1-k,\ui}$ and $\tilde P_{m+1-k,\ui'}$ are in fact equal, however $\tilde P_{m+1-k,\ui}$ is naturally identified as a subrepresentation of $\tilde P_{m+1}$ while $\tilde P_{m+1-k,\ui'}$ is not.
Indeed, by considering the support quiver of $\tilde P_{3,i}$ from Example~\ref{ex:lifted preprojectives}, we see that each $\tilde P_{2,(i,i)}$ is just a copy of $\tilde P_2$ when viewed as representations of $\widetilde{K(n)}$, however these provide distinct subrepresentations of $\tilde P_4$.
\begin{lemma}
  \label{le:special subrepresentations}
  Consider a nonempty subset $I\subsetneq\{1,\ldots,n\}$ and fix an element $j\in I$. 
  \begin{enumerate}
    \item For $m\ge2$, we have $\Hom(\tilde P_{m,j},\tau\tilde P_{m+1}^I)\cong\CC$.
      Moreover, the kernel of a nonzero morphism $\tilde P_{m,j}\to\tau\tilde P_{m+1}^I$ is the following representation 
      \[\tilde P_m(I,j):=
        \begin{cases}
          \bigoplus_{\substack{1\leq i\leq n\\i\neq j}}\tilde P_{m-1,(j,i)}\oplus \bigoplus_{i\in I, i\ne j}\tilde P_{m-2,(j,j,i)} & \text{ if $m\geq 3$;}\\
          \bigoplus_{\substack{1\leq i\leq n\\i\neq j}}\tilde P_{1,(j,i)} & \text{ if $m=2$.}
        \end{cases}\]
    \item For $m\ge3$, any nonzero morphism $\tilde P_{m,j}\to\tau\tilde P_{m+1}^I$ is surjective.
  \end{enumerate}
\end{lemma}
\begin{proof}
  The first claim of part (1) follows immediately by applying the Auslander-Reiten formulas \cite[Theorem IV.2.13]{ass} to Lemma~\ref{le:properties}.(3).
  Indeed, this gives
  \[\dim\Hom(\tilde P_{m,j},\tau\tilde P_{m+1}^I)=\dim\Ext(\tilde P_{m+1}^I,\tilde P_{m,j})=1.\]
  We establish the final claim of part (1) directly for $m=2,3$ and then deduce the general case by applying the reflection recursions \eqref{eq:recursive covers}.
  Using the description in Example~\ref{ex:truncated lifts}, it is not hard to see that $\tau\tilde P_3^I$ is indecomposable with one-dimensional spaces at only the vertices $(1,e)$ and $(2,\alpha_i^{-1})$ for $i\in I$.
  But then for~$j\in I$, the image of the unique homomorphism $\tilde P_{2,j}\to\tau\tilde P_3^I$ is the representation with support quiver
  \[(2,\alpha_j^{-1})\xrightarrow{\alpha_j}(1,e).\]
  From the support quiver of $\tilde P_{2,j}$ given in Example~\ref{ex:lifted preprojectives}, we see that the kernel of this map is precisely $\tilde P_2(I,j)$.

@@

The following orthogonality property is a primary reason we need to lift to the universal cover of $K(n)$.
\begin{corollary}
  \label{cor:perpendicular}
  Consider a nonempty subset $I\subsetneq\{1,\ldots,n\}$ and fix an element $j\in I$.
  For $m\geq 2$, we have $\tilde P_m(I,j)\in (\tilde P_{m+1}^I)^\perp$.
\end{corollary}
\begin{proof}
  If $m\geq 3$, we consider the long exact sequence obtained when applying $\Hom(\tilde P_{m+1}^I,-)$ to the sequence
  \[\ses{\tilde P_m(I,j)}{\tilde P_{m,j}}{\tau\tilde P_{m+1}^I}.\]
  Following Lemma~\ref{le:properties}.(3), we have $\Hom(\tilde P_{m+1}^I,\tilde P_{m,j})=0$ and using the long exact sequence this immediately implies $\Hom\big(\tilde P_{m+1}^I,\tilde P_m(I,j)\big)=0$.
  From Lemma~\ref{le:properties}.(3) again, we have $\Ext(\tilde P_{m+1}^I,\tilde P_{m,j})\cong\CC$.
  Using the Auslander-Reiten formulas \cite[Theorem IV.2.13]{ass} and Lemma~\ref{le:properties}.(4), we get
  \[\dim\Ext(\tilde P_{m+1}^I,\tau\tilde P_{m+1}^I)=\dim\Hom(\tilde P_{m+1}^I,\tilde P_{m+1}^I)=1.\]
  It follows that the surjective map $\Ext(\tilde P_{m+1}^I,\tilde P_{m,j})\to\Ext(\tilde P_{m+1}^I,\tau\tilde P_{m+1}^I)$ appearing in the long exact sequence is in fact an isomorphism.
  But the Auslander-Reiten formulas and Lemma~\ref{le:properties}.(4) again imply 
  \[\dim\Hom(\tilde P_{m+1}^I,\tau\tilde P_{m+1}^I)=\dim\Ext(\tilde P_{m+1}^I,\tilde P_{m+1}^I)=0.\]
  Combining with the preceding discussion, this gives $\Ext\!\big(\tilde P_{m+1}^I,\tilde P_m(I,j)\big)=0$.

@@

\subsection{Bialynicki-Birula decomposition}
The aim of Section \ref{torusaction} is to define a torus action on quiver Grassmannians which can be used to simplify the calculation of homological invariants in general.
If the quiver Grassmannian is smooth, which is for instance the case for exceptional representations by \cite{cr}, it can also be used to stratify the quiver Grassmannians using the results of Bialynicki-Birula.
More specifically, let $X$ be a smooth projective variety with a $\C^\ast$-action.
For a connected component of the fixed point set $C\subset X^{\C^\ast}$, we define its attracting set as
\[\mathrm{Att}(C):=\{y\in X\mid \lim_{t\to 0}t.y\in C\}\]
The following result of Bialynicki-Birula relates the geometry of $X$ to the geometry of its torus fixed points.
\begin{theorem}\label{thm:bb}
  \cite[Section 4]{bb}
  Let $X^{\C^\ast}=\coprod_{i=1}^r C_i$ be the decomposition of the fixed point set of $X$ into connected components.
  Then $\mathrm{Att}(C_i)$ is a locally closed smooth $\C^\ast$-invariant subvariety of $X$ containing $C_i$ as a subvariety.
  Moreover, we have $X=\coprod_{i=1}^r\mathrm{Att}(C_i)$ and the natural map $\gamma_i:\mathrm{Att}(C_i)\to C_i$ is an affine vector bundle. 
  In particular, we have an equality of Euler characteristics $\chi(X)=\chi(X^{\C^\ast})$.
\end{theorem}

@@

We can consider the decomposition $X_i=\bigoplus_{\chi\in A_Q} X_{(i,\chi)}$ and define a torus action on each $X_{(i,\chi)}$ via $t.x_{(i,\chi)}=t^{d(i,\chi)}x_{(i,\chi)}$ which can be extended linearly to each $X_i$.
Associated to each subspace $U_i$, there is a corresponding subspace $t.U_i$ for each $t\in\CC^*$.
In general, this does not induce a torus action on the quiver Grassmannians $\Gr_\bfe(X)$ since $t.U=(t.U_i)_{i\in Q_0}$ is not necessarily a subrepresentation of $X$ for every $U\in\Gr_\bfe(X)$.
Indeed, for this such an action must satisfy $X_\alpha(t.U_i)\subset t.U_j$ for every arrow $\alpha:i\to j$ of $Q$. 

@@

  Let $X$ be a representation of $Q$ which can be lifted to $\hat Q$.
  Fix an integer $c_\alpha\in\ZZ$ for each~$\alpha\in Q_1$.
  Suppose $d:\hat Q_0\to\ZZ$ satisfies $d(j,\chi+e_\alpha)-d(i,\chi)=c_\alpha$ for each arrow $\alpha:i\to j$ of $Q$ and each $\chi\in A_Q$.
  Then the torus action on $X$ determined by $d$ induces a torus action on $\Gr_\bfe(X)$.

@@

  Since $X$ can be lifted, for any arrow $\alpha\in Q_1$ we can write $X_\alpha:X_i\to X_j$ as a block matrix consisting of linear maps $X_{(\alpha,\chi)}:X_{(i,\chi)}\to X_{(j,\chi+e_\alpha)}$ for $\chi\in A_Q$.
  Fix $U\in\Gr_\bfe(X)$ and consider $u_i\in U_i$.
  Then we can write $u_i=\sum_{\chi\in A_Q} u_{(i,\chi)}$ for some vectors $u_{(i,\chi)}\in U_i\cap X_{(i,\chi)}$.
  But observe that for an arrow $\alpha:i\to j$ of $Q$, we have $X_\alpha(u_{(i,\chi)})=X_{(\alpha,\chi)}(u_{(i,\chi)})\in U_j\cap X_{(j,\chi+e_\alpha)}$, say $X_{(\alpha,\chi)}(u_{(i,\chi)})=u_{(j,\chi+e_\alpha)}$.
  It follows that
  \begin{align*}
    X_\alpha(t.u_i)=\sum\limits_{\chi\in A_Q}X_{(\alpha,\chi)}(t.u_{(i,\chi)})=\sum\limits_{\chi\in A_Q} t^{d(i,\chi)}X_{(\alpha,\chi)}(u_{(i,\chi)})=\sum\limits_{\chi\in A_Q} t^{d(i,\chi)}u_{(j,\chi+e_\alpha)}=t^{-c_\alpha}\sum\limits_{\chi\in A_Q} t.u_{(j,\chi+e_\alpha)}
  \end{align*}
  and thus $X_\alpha(t.U_i)\subset t.U_j$ for every arrow $\alpha:i\to j$ of $Q$.

@@

\begin{lemma}
  \label{degreecondition} 
  Let $X$ be an indecomposable representation of $\hat Q$.
  There exists $d:\mathrm{supp}(X)\to\ZZ$ and $c_\alpha\in\mathbb N_+$ for each $\alpha\in Q_1$ such that
  \begin{enumerate}
    \item for $(i,\chi),\,(i,\chi')\in\mathrm{supp}(X)$ with $\chi\neq\chi'$, we have $d(i,\chi)\ne d(i,\chi')$;
    \item for $(i,\chi),\,(j,\chi')\in\mathrm{supp}(X)$, we have $d(j,\chi')-d(i,\chi)=c_\alpha$ if and only if $\chi'=\chi+e_\alpha$.
  \end{enumerate}
\end{lemma}
\begin{proof}
  For convenience we introduce the notation $Q_1=\{\alpha_1,\ldots,\alpha_\}$.
  Since $X$ is finite-dimensional and indecomposable, the support quiver $\mathrm{supp}(X)$ is a connected and finite subquiver of $\hat Q$.
  In order to prove the statement, we may assume that $(i,0)\in\mathrm{supp}(X)$ for some $i\in Q_0$.
  Let $K$ be the maximal length of a path in $\mathrm{supp}(X)$ starting or ending in $(i,0)$ such that the underlying graph of the path has no cycles.
  This implies that, for $(i',\chi)\in\mathrm{supp}(X)$ with $\chi=\sum_{l=1}^n \kappa_l e_{\alpha_l}$, we have $|\kappa_l|\leq K$.

  Set $c_{\alpha_1}=1$ and choose $c_{\alpha_l}$ recursively in such way that 
  \[c_{\alpha_l}\geq 2(K+1)\sum_{k=1}^{l-1} c_{\alpha_k}\]
  for $l=2,\ldots,n$.

  Let $f:A_Q\to\ZZ$ be the group homomorphism defined by $f(e_\alpha)=c_\alpha$ for all $\alpha\in Q_1$ and define $d(i',\chi):=f(\chi)$ for $(i',\chi)\in\mathrm{supp}(X)$.
  In particular, for $(i',\chi)$ with $\chi=\sum_{l=1}^n \kappa_l e_{\alpha_l}$, we have
  \[d(i',\chi)=\sum_{l=1}^n \kappa_l c_{\alpha_l}.\]

  Now assume that $d(i',\chi)=d(i',\chi')$ for $\chi=\sum_{l=1}^n \kappa_l e_{\alpha_l}$ and $\chi'=\sum_{l=1}^n \kappa'_l e_{\alpha_l}$.
  This implies
  \[\sum_{l=1}^{n-1}(\kappa_l-\kappa'_l)c_{\alpha_l}=(\kappa'_n-\kappa_n)c_{\alpha_n}.\]
  But we have $|\kappa_l-\kappa'_l|\leq |\kappa_l|+|\kappa'_l|\le 2K$ and thus we obtain 
  \[|\sum_{l=1}^{n-1}(\kappa_l-\kappa'_l)c_{\alpha_l}|\le 2K\sum_{l=1}^{n-1}c_{\alpha_l} < c_{\alpha_n}.\]
  This inductively yields $\kappa_l=\kappa'_l$ for $l=n,\ldots,1$ by the choice of the $c_{\alpha_l}$.

  By definition, we have $d(j,\chi+e_\alpha)-d(i,\chi)=c_\alpha$.
  Assume that $d(j,\chi')-d(i,\chi)=c_\alpha$.
  Then we have $d(j,\chi')=d(j,\chi+e_\alpha)$.
  An analogous proof as the one above shows that $\chi'=\chi+e_\alpha$.
  Note that it is a priori possible that $(j,\chi+e_\alpha)\notin\mathrm{supp}(X)$.
  Thus we chose $c_{\alpha_j}$ in such a way that it is greater than $2(K+1)\sum_{i=1}^{j-1}c_{\alpha_i}$. 
\end{proof}
